\documentclass[11pt]{article}
\input{_common_preamble}
\input{_common_notation}

\title{Phase 1 Polymer Online Re-Identification With Batch Ridge And RL Blend}
\date{2026-04-12}

\begin{document}
\maketitle

\section*{Method Goal}

This method adds a polymer-only online re-identification workflow on top of the
current unified notebook surface. The observer remains nominal and fixed, while
the MPC prediction model is blended online between the nominal model and a
rolling identified model:
\[
A_{\mathrm{pred},k} = A_0 + \eta_k (A_{\mathrm{id},k} - A_0), \qquad
B_{\mathrm{pred},k} = B_0 + \eta_k (B_{\mathrm{id},k} - B_0).
\]

The RL policy does not act directly on model multipliers. It chooses a scalar
blend factor $\eta_k \in [0,1]$.

\section*{Notebook Source}

\begin{itemize}
\item \texttt{RL\_assisted\_MPC\_reid\_batch\_unified.ipynb}
\item Shared helpers:
      \texttt{utils/reid\_batch.py},
      \texttt{utils/reid\_batch\_runner.py},
      \texttt{utils/plotting.py},
      \texttt{utils/plotting\_core.py}
\end{itemize}

\section*{State, Action, And Reward}

The base RL state uses the current unified state builder in either standard or
mismatch mode:
\[
s_t^{\mathrm{base}} = \texttt{build\_rl\_state}(\hat{x}_t, y_{\mathrm{sp},t}, u_t).
\]

The blend-policy state appends online-ID diagnostics:
\[
s_t =
\begin{bmatrix}
s_t^{\mathrm{base}} \\
\eta_{t-1} \\
\|\varepsilon_t^{\mathrm{id}}\|_2 \\
\|\Delta A_t\|_F / \|A_0\|_F \\
\|\Delta B_t\|_F / \|B_0\|_F \\
\mathbb{1}_{\mathrm{id\_valid}}
\end{bmatrix}.
\]

The RL action is one scalar:
\[
a_t \in [-1, 1], \qquad
\eta_t^{\mathrm{raw}} = \frac{a_t + 1}{2}, \qquad
\eta_t = (1-\tau_\eta)\eta_{t-1} + \tau_\eta \eta_t^{\mathrm{raw}}.
\]

The reward is the same MPC-style stage reward already used in the unified
polymer notebook family.

\section*{Identification Model}

Phase 1 reuses the current polymer scalar-matrix basis because it already fits
the repo's linear-in-parameter workflow:

\begin{itemize}
\item one global physical-$A$ correction parameter
\item one physical-$B$ correction parameter per input column
\end{itemize}

The batch regression is assembled on physical scaled-deviation dynamics:
\[
x_{t+1}^{\mathrm{phys}} - A_0 x_t^{\mathrm{phys}} - B_0 u_t
= \Phi_t \theta.
\]

The default solver is ridge around the previous estimate:
\[
(\Phi^\top \Phi + (\lambda_{\mathrm{prev}} + \lambda_0)I)\theta
= \Phi^\top r + \lambda_{\mathrm{prev}}\theta_{\mathrm{prev}}.
\]

An optional bounded least-squares mode is also supported through
\texttt{scipy.optimize.lsq\_linear}.

\section*{Runtime Flow}

\begin{enumerate}
\item Build the nominal observer state and the current blend-policy state.
\item Choose the scalar RL action and map it to $\eta_t$.
\item Form the blended prediction model from the current identified model and
      $\eta_t$.
\item Solve MPC using the blended prediction model.
\item Step the plant.
\item Update the observer using only the nominal augmented model and fixed
      observer gain.
\item Push the physical transition into the rolling identification buffer.
\item Every \texttt{id\_update\_period} samples, solve the batch ID problem on
      the last \texttt{id\_window} samples and accept the candidate only if the
      identified model remains within configured $\Delta A$ / $\Delta B$ caps.
\item Store the transition for RL and train the chosen continuous agent.
\end{enumerate}

\section*{Defaults Chosen}

\begin{itemize}
\item polymer only in phase 1
\item default agent: TD3
\item \texttt{id\_solver = ridge\_closed\_form}
\item \texttt{id\_window = 80}
\item \texttt{id\_update\_period = 5}
\item \texttt{lambda\_prev = 10^{-2}}
\item \texttt{lambda\_0 = 10^{-4}}
\item \texttt{eta\_smoothing\_tau = 0.1}
\item $\eta = 0$ during warm start
\end{itemize}

\section*{Diagnostics}

The saved bundle records:

\begin{itemize}
\item $\eta$ and raw-action traces
\item theta history
\item identification residual and conditioning traces
\item identified-model and blended-model Frobenius delta ratios
\item ID update event / success / fallback flags
\item standard closed-loop outputs, inputs, rewards, and observer traces
\end{itemize}

\section*{Scope Limits}

\begin{itemize}
\item Phase 1 is polymer only.
\item The observer remains nominal and fixed.
\item No changes are made to the existing matrix, structured-matrix, baseline,
      horizon, weights, residual, or combined workflows.
\item The richer row-wise identification basis is deferred to a later phase.
\end{itemize}

\end{document}
